---
title: "GLM MCMC diagnostics"
author: "Katy Gaythorpe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes: 
- \usepackage{placeins}
output: 
  pdf_document:
    df_print: "kable"
---

```{r knitsetup, include=FALSE}

knitr::opts_chunk$set(
  dev = "png",
  fig.path = "images/",
  dpi = 300,
  warning = FALSE,
  message = FALSE
)
```

Here, we process the output of the GLM estimation.

```{r setup}
library(mcmcplots)
library(ggmcmc)
library(gridExtra)
library(maptools)
library(readr)
library(fields)
library(rgdal)
library(R.utils)
library(magrittr)
library(pROC)
library(dplyr)

# homemade
library(snapalette)
library(YFestimation)
library(KsetupR)

R.utils::sourceDirectory("Functions")

# glm_mcmc_out <- YFestimation::get_chains("GLM_MCMC_chain_20190919_6_new_pop_dat_wes",
#                          burnin =1e5, thin = 1)

glm_mcmc_out <- YFestimation::get_chains("GLM_MCMC_chain_20190919_6_new_pop_dat_wes_noadm05",
                          burnin =1e4, thin = 1)

names(glm_mcmc_out) <- gsub("^log.", "", names(glm_mcmc_out))

glm_mcmc_out2 <- ggs( convert.mcmc.list(  glm_mcmc_out %>% 
                                           dplyr::select(-c(posteriorProb, 
                                                     acceptRate, 
                                                     starts_with("adm05")))) )

snapal <- "Pop"
```

First we plot the posterior.

```{r posterior,   fig.cap = "Posterior.", fig.align='center', fig.pos = 'h', strip.white=F }
plot(glm_mcmc_out$posteriorProb, 
     type = "l", 
     xlab = "Iteration", 
     ylab = "Posterior probability")

maxind <- which.max(glm_mcmc_out$posteriorProb )

#write.csv(glm_mcmc_out[maxind,], "glm_param.csv", row.names = FALSE)
```


To assess the convergence of the MCMC chains we shall plot a selection of the outputs for two parameters.

```{r trace_glm,   fig.cap = "\\label{fig:trace_glm}Trace plot for each parameter.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10 }

cols <- snapalette(snapal, ncol(glm_mcmc_out)-2, type = "continuous")
par(mfrow = c(4,3))
for(i in 1:(ncol(glm_mcmc_out)-2)){
  plot(glm_mcmc_out[, i], type = "l", main = names(glm_mcmc_out)[i], col = cols[i])
}

```

\FloatBarrier

```{r density_glm, fig.cap = "\\label{fig:density_glm}Density plot for each parameter.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10}
par(mfrow = c(4,3))
for(i in 1:(ncol(glm_mcmc_out)-2)){
  
  plot(density(glm_mcmc_out[, i]), main = names(glm_mcmc_out)[i], col = cols[i])
  polygon(density(glm_mcmc_out[, i]), col = cols[i])
}

```
\FloatBarrier

Finally, we examine the values of all parameters through a caterpillar plot and compare the prior and posterior distributions.

```{r caterpillar_glm,   fig.cap = "\\label{fig:caterpillar_glm}Caterpillar plot of parameters excluding adm05.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10}

ggs_caterpillar(glm_mcmc_out2)
```

```{r crosscor_glm,   fig.cap = "Crosscorrelation of all parameters excluding adm05.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10}

ggs_crosscorrelation(glm_mcmc_out2)
```

```{r priorpost_glm,   fig.cap = "\\label{fig:priorpost_glm}Prior (yellow) and posterior (grey) distributions for all parameters.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10, eval = FALSE}

plot_prior_post2(glm_mcmc_out, 
                prior_col = snapalette(snapal, 5, type="discrete")[5], 
                "GLM")
```

\FloatBarrier 

## Assessing the fit

We now collect the estimates to produce the posterior predictive distribution of yellow fever reports.

```{r glm_predict,  fig.cap = "Prediction of GLM including surveillance quality parameters. The scale indicates the probability of YF report over the observation period.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10, fig.width=10}


Est_beta <-  apply(subset( glm_mcmc_out, 
                          select = -c(posteriorProb, acceptRate)), 
                  2, median)


### LOADING SHAPEFILES AND COUNTRIES ###
#read shapefiles in
shp0 <- rgdal::readOGR("../../shapefiles/Africa_America_0_smooth.shp",
                          stringsAsFactors = FALSE,
                          encoding = "UTF-8",
                          use_iconv = TRUE)
shp1 <- rgdal::readOGR("../../shapefiles/Africa_America_1_smooth.shp",
                          stringsAsFactors = FALSE,
                          encoding = "UTF-8",
                          use_iconv = TRUE)


#sort age vector
maxAge <- 100
ageVec <- c(0:maxAge)


### LOAD ENVIRONMENTAL DATA ###
Env_Table_path <- "../Data/Environment/global_dat"

filename <- KsetupR::get_latest_file(path = Env_Table_path, pattern = "dat_wes")

dat <- read.csv(filename, stringsAsFactors = FALSE)

# remove families of NHP that are not to be included
model_form_whole <- read.csv("Model_form6.csv", stringsAsFactors = FALSE)$x
covar <- unlist(strsplit(unlist(strsplit(model_form_whole, "\\+")), "\\~"))

dat <- dat[, -which(!names(dat) %in% covar & grepl("family", names(dat)))] 

dat %<>% dplyr::select(-year)
dat %<>% filter( is.finite(MIR.max))
dat %<>% filter(!is.na(precip_mean))

# make extra elements and normalise
dat = adjust_env_dat(dat)
# -----------------------------------------------------------------------
# FIT MODEL #
covar <- covar[grep("cases_or_outbreaks|family", covar, invert = TRUE)]

model_form <- paste0("cases_or_outbreaks~", paste(covar, collapse = "+") , "+aggregate_family")#,  "+adm05", "+surv.qual.adm0")

object_glm <- YFestimation::fit_glm(dat = dat, 
                                   depi = match("cases_or_outbreaks", 
                                                names(dat)), 
                                   models = model_form)

beta0 <- object_glm[[1]]
x <- object_glm[[2]]
y <- object_glm[[3]]

#order Est_beta to match beta0
Est_beta = Est_beta[names(beta0)]

###########################################################################
plot_glm_map2(shp0, shp1,  dat, Est_beta, x, 
              colours = snapalette(snapal, 1000, type = "continuous"), 
              plot_data = FALSE)

```

```{r AUC}

glmpreds <- YFestimation::fun_calcPred( Est_beta ,x,type="response")

ROC_out <- pROC::roc(response = dat$cases_or_outbreaks, predictor = glmpreds)

auc(ROC_out)

```

```{r plot_wo_adm05, fig.cap = "Prediction of GLM excluding surveillance. The scale indicates the probability of YF occurrence over the observation period.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10, fig.width=10}

x_wo_adm05 <- x[, grep("adm05|surv", colnames(x), invert = TRUE)]
Est_beta_wo_adm05 <- Est_beta[grep("adm05|surv", names(Est_beta), 
                                  invert = TRUE)]

colours <- snapalette::snapalette(snapal, 1000, type = "continuous")

eta <-  x_wo_adm05 %*% Est_beta_wo_adm05
preds <- 1-exp(-exp(eta)) # 

shp1$predictions  <- preds[match( shp1$GID_1, dat$adm1)]

mybreaks <- seq(0, 1.0001, length.out=1001)
mycols <-  colours
vcols <- findInterval(shp1$predictions, mybreaks)


plot(shp0)
plot(shp1, col=mycols[vcols] , lty=0, add=TRUE)
plot(shp0, add=TRUE)
fields::image.plot(legend.only=TRUE, 
                   breaks=mybreaks, 
                   col=mycols, 
                   zlim=c(0,1), 
                   horizontal = TRUE)

```

What proportion of areas are predicted correctly?

```{r proportion_correct}

df = data.frame(shp1$GID_0, shp1$GID_1, shp1$predictions, dat$cases_or_outbreaks[match(shp1$GID_1, dat$adm1)])
names(df) = c("adm0", "adm1", "predictions", "data")

df1 = df %>% filter(data == 0) %>% filter(predictions <=0.5)
df2 = df %>% filter(data == 1) %>% filter(predictions >=0.5)

df3 = data.frame("x1" = nrow(df1)/nrow(df %>% filter(data == 0)),
           "X2" = nrow(df2)/nrow(df %>% filter(data == 1)))

names(df3) = c("Proportion of negatives correctly predicted", "Proportion of positives correctly predicted")
df3
```


```{r rayshader_prep, eval = F, echo = F}
library(rayshader)
library(raster)

dim_out = 1000
 

blank_raster <- raster(shp1, ncols = dim_out, nrows = dim_out, vals = 0)
each_adm_value <- sapply(1:length(shp1$predictions), function(i){
  #print(i)
  intersect_country <- rasterize(shp1[i, ], crop(blank_raster, extent(shp1[i, ])), mask = TRUE)
  intersect_country[!is.na(intersect_country)] <- shp1$predictions[i]
  intersect_country

}, simplify = FALSE)

each_adm_value$fun <- mean

full_shape <- do.call(mosaic, each_adm_value)

plot(full_shape)
```

```{r rayshader_out, eval = F, echo = F}

bath = focal(flip(full_shape, "x"), w=matrix(1,3,3), mean)

tempfilename = tempfile()
png(tempfilename,width = 401,height=401)
par(mar = c(0,0,0,0))
raster::image(bath,axes = FALSE, col = snapalette::snapalette(snapal, 1000, type = "continuous") )
dev.off()
tempmap = png::readPNG(tempfilename)

bath[is.na(bath)] = -1
depth_exaggeration = 100

bath %>%
  as.matrix() %>%
  sphere_shade(texture="desert",
               zscale = 1/depth_exaggeration) %>%
  add_shadow(ray_shade(as.matrix(bath), 
                       zscale = 1/depth_exaggeration), 
             max_darken = 1) %>%
  add_shadow(ambient_shade(as.matrix(bath), 
                           zscale = 1/depth_exaggeration)) %>%
  
  add_overlay(tempmap, alphacolor = 0.5) %>%

  plot_3d(as.matrix(bath),
          solid=TRUE,
          soliddepth="auto",
          solidcolor = 'white',
          solidlinecolor = "grey50",
          water = TRUE,
          waterdepth = 0,
          wateralpha = 0.9,
          theta = 90,
          shadow = TRUE,
          background='white',
          zscale = 1/depth_exaggeration,
          windowsize = c(1200, 700)) 

render_snapshot(clear = TRUE)

```


