---
title: "GLM MCMC diagnostics"
author: "Katy Gaythorpe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes: 
- \usepackage{placeins}
output: 
  pdf_document:
    df_print: "kable"
params:
  model_variant: 
    value: 1
---

```{r knitsetup, include=FALSE}

knitr::opts_chunk$set(
  dev = "png",
  fig.path = paste0("images", params$model_variant, "/"),
  dpi = 300,
  warning = FALSE,
  message = FALSE
)
```

Here, we process the output of the GLM estimation.

```{r setup}
library(mcmcplots)
library(ggmcmc)
library(gridExtra)
library(maptools)
library(readr)
library(fields)
library(rgdal)
library(R.utils)
library(magrittr)
library(pROC)
library(dplyr)

# homemade
library(snapalette)
library(YFestimation)
library(KsetupR)

R.utils::sourceDirectory("Functions")


glm_mcmc_out <- YFestimation::get_chains(paste0("GLM_MCMC_chain_20191016_bestglm_1_", 
                                                params$model_variant),
                          burnin =6e4, thin = 1)


names(glm_mcmc_out) <- gsub("^log.", "", names(glm_mcmc_out))

glm_mcmc_out2 <- ggs( convert.mcmc.list(  glm_mcmc_out %>% 
                                           dplyr::select(-c(posteriorProb, 
                                                     acceptRate))))

snapal <- "Pop"
```

First we plot the posterior.

```{r posterior,   fig.cap = "Posterior.", fig.align='center', fig.pos = 'h', strip.white=F }
plot(glm_mcmc_out$posteriorProb, 
     type = "l", 
     xlab = "Iteration", 
     ylab = "Posterior probability")

maxind <- which.max(glm_mcmc_out$posteriorProb )

#write.csv(glm_mcmc_out[maxind,], "glm_param.csv", row.names = FALSE)
```


To assess the convergence of the MCMC chains we shall plot a selection of the outputs for two parameters.

```{r trace_glm,   fig.cap = "\\label{fig:trace_glm}Trace plot for each parameter.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10 }

cols <- snapalette(snapal, ncol(glm_mcmc_out)-2, type = "continuous")
par(mfrow = c(4,3))
for(i in 1:(ncol(glm_mcmc_out)-2)){
  plot(glm_mcmc_out[, i], type = "l", main = names(glm_mcmc_out)[i], col = cols[i])
}

```

\FloatBarrier

```{r density_glm, fig.cap = "\\label{fig:density_glm}Density plot for each parameter.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10}
par(mfrow = c(4,3))
for(i in 1:(ncol(glm_mcmc_out)-2)){
  
  plot(density(glm_mcmc_out[, i]), main = names(glm_mcmc_out)[i], col = cols[i])
  polygon(density(glm_mcmc_out[, i]), col = cols[i])
}

```
\FloatBarrier

Finally, we examine the values of all parameters through a caterpillar plot and compare the prior and posterior distributions.

```{r caterpillar_glm,   fig.cap = "\\label{fig:caterpillar_glm}Caterpillar plot of parameters excluding adm05.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10}

ggs_caterpillar(glm_mcmc_out2)
```

```{r crosscor_glm,   fig.cap = "Crosscorrelation of all parameters excluding adm05.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10}

ggs_crosscorrelation(glm_mcmc_out2)
```

```{r continent_correlation, eval = F}

ggplot(glm_mcmc_out, aes(x = adm05BRA, y = continent_factor)) + 
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon")+
     scale_fill_gradient(low="blue", high="red") + 
  theme_bw()
```

```{r acceptance}
mean(glm_mcmc_out$acceptRate)

```



```{r priorpost_glm,   fig.cap = "\\label{fig:priorpost_glm}Prior (yellow) and posterior (grey) distributions for all parameters.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10, eval = FALSE}

plot_prior_post2(glm_mcmc_out, 
                prior_col = snapalette(snapal, 5, type="discrete")[5], 
                "GLM")
```

\FloatBarrier 

## Assessing the fit

We now collect the estimates to produce the posterior predictive distribution of yellow fever reports.

```{r glm_predict,  fig.cap = "Prediction of GLM including surveillance quality parameters. The scale indicates the probability of YF report over the observation period.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 8, fig.width=10}


Est_beta <-  apply(subset( glm_mcmc_out, 
                          select = -c(posteriorProb, acceptRate)), 
                  2, median)


### LOADING SHAPEFILES AND COUNTRIES ###
#read shapefiles in
shp0 <- rgdal::readOGR("../../shapefiles/Africa_America_0_smooth.shp",
                          stringsAsFactors = FALSE,
                          encoding = "UTF-8",
                          use_iconv = TRUE)
shp1 <- rgdal::readOGR("../../shapefiles/Africa_America_1_smooth.shp",
                          stringsAsFactors = FALSE,
                          encoding = "UTF-8",
                          use_iconv = TRUE)


#sort age vector
maxAge <- 100
ageVec <- c(0:maxAge)


### LOAD ENVIRONMENTAL DATA ###
Env_Table_path <- "../Data/Environment/global_dat"

filename <- KsetupR::get_latest_file(path = Env_Table_path, pattern = "dat_wes")

dat <- read.csv(filename, stringsAsFactors = FALSE)

#remove families of NHP that are not to be included
model_form_whole = read.csv("bestglm_1.csv", stringsAsFactors = FALSE) %>% 
  dplyr::select(-Criterion)

covar_family = names(model_form_whole)[ which(model_form_whole[params$model_variant, ] == TRUE) ] 

dat <- dat[, -which(!names(dat) %in% covar_family & grepl("family", names(dat)))] 

# change to just read covariates from mcmc
covar = names(glm_mcmc_out %>% dplyr::select(-c(Intercept, 
                                                starts_with("adm05"), 
                                                posteriorProb, 
                                                acceptRate)))

  
    
dat %<>% dplyr::select(-year)
dat %<>% filter( is.finite(MIR.max))
dat %<>% filter(!is.na(temp_mean))
#dat %<>% tidyr::drop_na()
dat %<>% mutate(continent = ifelse(adm0 %in% c("CIV", "STP"),
                                   "Africa", continent))

#dat %<>% filter(risk %in% c("high", "moderate"))

# make extra elements and normalise
dat = adjust_env_dat(dat)

#dat %<>% filter(continent == "Africa")
# c34 = readr::read_csv("../Data/Countries.csv")$c34
# dat %<>% filter(adm0 %in% c34)

# -----------------------------------------------------------------------
# FIT MODEL #

model_form <- paste0("cases_or_outbreaks~", paste(covar, collapse = "+") , "+adm05") 

as.data.frame(c(covar))#, "adm05"))
object_glm <- YFestimation::fit_glm(dat = dat, 
                                   depi = match("cases_or_outbreaks", 
                                                names(dat)), 
                                   models = model_form)

beta0 <- object_glm[[1]]
x <- object_glm[[2]]
y <- object_glm[[3]]

#order Est_beta to match beta0
Est_beta = na.omit(Est_beta[names(beta0)])

###########################################################################
plot_glm_map2(shp0, shp1,  dat, Est_beta, x, 
              colours = snapalette(snapal, 1000, type = "continuous"), 
              plot_data = FALSE)

```

```{r AUC}

glmpreds <- YFestimation::fun_calcPred( Est_beta ,x,type="response")

ROC_out <- pROC::roc(response = dat$cases_or_outbreaks, predictor = glmpreds)

print(paste0("All ", round(auc(ROC_out), 4)))

if(length(unique(dat$continent))>1){
  
  africa_ind = which(dat$continent == "Africa")
  america_ind = which(dat$continent != "Africa")
  
  ROC_out <- pROC::roc(response = dat$cases_or_outbreaks[africa_ind], predictor = glmpreds[africa_ind])
  print(paste0("Africa ", round(auc(ROC_out), 4)))
  
  ROC_out <- pROC::roc(response = dat$cases_or_outbreaks[america_ind], predictor = glmpreds[america_ind])
  print(paste0("South America ", round(auc(ROC_out), 4)))
}

```

```{r plot_wo_adm05, fig.cap = "Prediction of GLM excluding surveillance. The scale indicates the probability of YF occurrence over the observation period.", fig.align='center', fig.pos = 'h', strip.white=F, fig.height = 10, fig.width=10}

x_wo_adm05 <- x[, grep("adm05|surv|dtp", colnames(x), invert = TRUE)]
Est_beta_wo_adm05 <- Est_beta[grep("adm05|surv|dtp", names(Est_beta), 
                                  invert = TRUE)]

colours <- snapalette::snapalette(snapal, 100, type = "continuous")

eta <-  x_wo_adm05 %*% Est_beta_wo_adm05
preds <- 1-exp(-exp(eta)) # 

shp1$predictions  <- preds[match( shp1$GID_1, dat$adm1)]

mybreaks <- seq(0, 1.0001, length.out=101)
mycols <-  colours
vcols <- findInterval(shp1$predictions, mybreaks)


plot(shp0)
plot(shp1, col=mycols[vcols] , lty=0, add=TRUE)
plot(shp0, add=TRUE)
fields::image.plot(legend.only=TRUE, 
                   breaks=mybreaks, 
                   col=mycols, 
                   zlim=c(0,1), 
                   horizontal = TRUE)

```

What proportion of areas are predicted correctly?

```{r proportion_correct}

df = data.frame(shp1$GID_0, 
                shp1$GID_1, 
                shp1$predictions, 
                dat$cases_or_outbreaks[match(shp1$GID_1, dat$adm1)])
names(df) = c("adm0", "adm1", "predictions", "data")

df1 = df %>% filter(data == 0) %>% filter(predictions <=0.5)
df2 = df %>% filter(data == 1) %>% filter(predictions >=0.5)

df3 = data.frame("x1" = nrow(df1)/nrow(df %>% filter(data == 0)),
           "X2" = nrow(df2)/nrow(df %>% filter(data == 1)))

names(df3) = c("Proportion of negatives correctly predicted", 
               "Proportion of positives correctly predicted")
df3
```



